---
title: "Statistical Analysis For Dataset Cities"
author: "Xinxin Xu USC ID 4205459366"
date: "2023-02-20"
output: pdf_document
---

```{r}
# import data
library(readxl)
df <- read_excel("cities1.xlsx")

library(cluster)  # silohuette()
library(factoextra)
#
str(df)
head(df)
```

```{r}
#drop character
df= as.data.frame(df)
df$Crime_Trend = NULL
df$Unemployment_Threat = NULL
row.names(df) = df$Metropolitan_Area
df$Metropolitan_Area = NULL
head(df,5)
```

```{r}
# scale data (always for distance-based methods)
df_scale = scale(df)
```  

```{r}
#  Use set.seed(123) and the user function twcv to find TWCV values for k = 1 : 16. Use nstart = 25. Display the elbow chart.

set.seed(123)
twcv = function(k) kmeans(df_scale, k, nstart = 25 )$tot.withinss
k <- 1:16
twcv_values <- sapply(k,twcv)
head(twcv_values)
plot(k, twcv_values,type="b",pch = 19, 
     xlab="Number of clusters K",ylab="TWCV")
grid()
```


```{r}
#The best number of clusters is the smallest k such that the cluster plot shows the least amount of clusters overlap. Use fviz_cluster( ) with argument geom = ”point” to display cluster plots with no label names. Try fviz_cluster( ) with different K. What is the best K? For this K find the number of cities in each cluster.
final6 <- kmeans(df_scale, center = 6, nstart = 25)
fviz_cluster(final6, data = df_scale, geom = 'point')

final4 <- kmeans(df_scale, center = 4, nstart = 25)
fviz_cluster(final4, data = df_scale, geom = 'point')

final3 <- kmeans(df_scale, center = 3, nstart = 25)
fviz_cluster(final3, data = df_scale, geom = 'point')

final2 <- kmeans(df_scale, center = 2, nstart = 25)
fviz_cluster(final2, data = df_scale, geom = 'point')

## I find the best value of K (the optimal number of clusters) is 4
cluster_number = final4$cluster
table(cluster_number)
length(cluster_number)
m1 = prcomp(df_scale, scale = T)
fviz_pca_biplot(m1, labelsize = 1, col.var = 'black', habillage = cluster_number)
```


```{r}
# I chose clustering with k = 4
#
set.seed(123)
final <- kmeans(df_scale,centers = 4,nstart = 25)
final

fviz_cluster(final,data = df_scale)
# add cluster number to dataframe
#
cluster_number = as.factor(final$cluster)
df$cluster = cluster_number
head(df)

# biplot with clusters
#
m3 = prcomp(df_scale, scale=T)
fviz_pca_biplot(m3,labelsize = 2,col.var = "red",
                habillage = cluster_number)
```



```{r}
# For my choice of K clusters, find the median (or mean, if you prefer) of each numerical column (on the original un-scaled dataset). 

df = lapply(df,as.numeric)
aggregate(df, list(cluster_number), median)

## Group (cluster) 1 & 2 have high rate on every category except cost of living and crime
## Group (cluster) 3 has high crime 
## Group (cluster) 4 has high rates on cost of living

```




```{r}

# Use function hclust with linkage ward.D to create object h1 and display the four clusters on the dendrogram. Use function cuttree( ) to find the clusters. Find the number of cities in each cluster. Use fviz_cluster( ) with argument geom = ”point” to display the cluster plots of your choice with no label names. Find the CCPC for ward.D

# Find distances
distance = dist(df_scale)

# Dendrogram - Ward
h1 = hclust(distance, method = 'ward.D')
plot(h1, cex = 0.4, xlab = '', main = 'ward.D', ylab = '')
rect.hclust(h1, k = 4, border = 'red')

# CUT the dendrograms to 4 clusters
cut1 = cutree(h1,k=4)
# dataframe with cluster numbers
df1 = data.frame(df,cluster = cut1)
# number of members per cluster
table(cut1)


# library factoextra
library(factoextra)

# cluster plots for Ward linkage
fviz_cluster(list(data = df_scale, cluster = cut1),geom = 'point', main="Ward linkage", ellipse.type = 'norm',
             palette = "Set2",
             ggtheme = theme_minimal()
             )

# CPCC = correlation (euclidean distances, cophenetic distances)
c1 = cophenetic(h1)
cor(distance, c1)
```

```{r}
# Use function hclust with linkage complete to create object h2 and display the four clusters on the dendrogram. Use function cuttree( ) to find the clusters. Find the number of cities in each cluster. Use fviz_cluster( ) with argument geom = ”point” to display the cluster plots of your choice with no label names. Find the CCPC for complete linkage.

h2 = hclust(distance, method ='complete')
cut2 = cutree(h2,k=4)
# number of members per cluster
table(cut2)

# dendrogram - complete linkage
plot(h2, cex = 0.4, xlab = '', main = 'Complete', ylab = '')
rect.hclust(h2, k = 4, border = 'red')

# cluster plots for complete linkage
fviz_cluster(list(data = df_scale, cluster = cut2), geom = 'point', 
              main = 'complete linkage',
              palatte = 'Set2', ggtheme = theme_minimal())


# COPHENETIC distances
c2 = cophenetic(h2)
# CPCC = correlation (euclidean distances, cophenetic distances)
cor(distance,c2)
```

```{r}

# Use function hclust with linkage average to create object h3 and display the four clusters on the dendrogram. Use function cuttree( ) to find the clusters. Find the number of cities in each cluster. Use fviz_cluster( ) with argument geom = ”point” to display the cluster plots of your choice with no label names. Find the CCPC for average linkage

# dendrogram - complete linkage
h3 = hclust(distance, method = 'average')
plot(h3, cex = 0.4, xlab = '', main = 'average', ylab = '')
rect.hclust(h3, k = 4, border = 'red')

# number of members per cluster
cut3 = cutree(h3, k = 4)
table(cut3)

# cluster plots for average linkage
fviz_cluster(list(data = df_scale, cluster = cut3), geom = 'point', 
             main = 'average', palatte = 'Set2', ggtheme = theme_minimal())


# CPCC = correlation (euclidean distances, cophenetic distances)
c3 = cophenetic(h3)
cor(distance, c3)
```

```{r}

# What linkage prefer? For the clusters found for this linkage find the median (or mean, if you prefer) of each numerical column (on the original un-scaled dataset). Write one sentence characterizing each cluster for this linkage.

print('I perfer ward.D because the corrlation is the smallest.')
aggregate(df, list(cut1), median)
```

## Group 1 have high rate on cost of living, but low on Health_Care and Recreation, maybe not a good choice if people want to relocate to area in this group.

## Group 2 have low rate on Cost_Living, Total_Property, but high rate on crime, indicating thus area is not safe, and very poor.

## Group 3 has high rate on climate, health care and total violent, but low on crime.

## Group 4 has high rates on education, arts, health care and recreation, population_2000, Past_Job_Growth, and Fcast_Future_Job_Growth, indicating group 4 as very suited for living. People there should live happily.
